Article data analysis

### 2018-10-08 notebooks/article_notebook.Rmd
updated with a new figure

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","aac","oqx")
mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")

@@

@@

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
heatmap_df <- create_heatmap_df(acquired_report)

@@

@@

mut_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_x_discrete(limits = c(
      "gyrA", "gyrB", "parC", "parE",
      "marR", "soxS", "tolC"
    )) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

>>>>

mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","oqx")
mut_genes <- c("gyr","par","marR","soxR","acrR")

@@

# Introduces unique ID's for each sample
fix_ref_id <- function(df) {
  df <- df %>%
    mutate(ref = gsub("(.*?)\\_.+", "\\1", ref),
         ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
         ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    select(-ref)
  return(df)
}

@@

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    mut_report_fixed <- fix_ref_id(mut_report)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
  mut_report_fixed <- fix_ref_id(mut_report)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
    acquired_report_fixed <- fix_ref_id(acquired_report)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
  acquired_report_fixed <- fix_ref_id(acquired_report)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
mut_quant_fixed <- fix_ref_id(mut_quant)
heatmap_df <- create_heatmap_df(acquired_report)

@@

## Percent isolates with PMQR and/or QRDR mutations in gyrA, gyrB, parC or parE
```{r}
total_report <- mut_report_fixed %>%
  left_join(acquired_report_fixed, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  select(-ref) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  group_by(present,CIP) %>%
  count() %>%
  ungroup() %>%
  rowwise() %>%
  mutate(present = ifelse(present == "", "None", present),
         total = 285,
         percent = round(n/total * 100, 1),
         lwr = round(get_binCI(n, total)[1], 1),
         upr = round(get_binCI(n, total)[2], 1))

genes <- c(names(mut_report_fixed), names(acquired_report_fixed))
genes <- genes[!genes == "id"]

p1 <- ggplot(total_report, aes(reorder(present, as.numeric(-CIP)), percent, fill = factor(CIP)))+
  geom_col(color = "black")+
  labs(y = "Percent (%) isolates")+
  scale_fill_viridis(discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_text(size = 10),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.position = "top",
        legend.spacing.x = unit(0.27, "cm"))

dot_report <- mut_report_fixed %>%
  left_join(acquired_report_fixed, by = "id") %>%
  gather(gene, value, -id) %>%
  mutate(present = ifelse(value == 1, gene, NA),
         ref = 1:n()) %>%
  spread(gene, value) %>%
  group_by(id) %>%
  summarise_all(funs(func_paste)) %>%
  left_join(mic_data[c("id","CIP")], by = "id") %>%
  select(-c(ref, id)) %>%
  ungroup() %>%
  mutate_at(vars(-present, CIP), .funs = as.numeric) %>%
  mutate(none = ifelse(rowSums(.[,genes]) == 0, 1, 0),
         present = ifelse(present == "", "None", present)) %>%
  gather(gene, value, -c(present,CIP)) %>%
  mutate(gene = ifelse(value == 0 & gene != "None", NA, gene),
         value = ifelse(gene == "None" & value == 0, NA, value)) %>%
  na.omit()

p2 <- ggplot(dot_report, aes(reorder(present, -CIP), gene, fill = gene))+
  geom_point(size = 2, pch = 21)+
  scale_fill_viridis(option = "B",
                     discrete = TRUE)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.title = element_blank(),
        axis.ticks.x = element_blank())

plot_grid(p1, p2, nrow = 2, align = "v", rel_heights = c(3.5/5, 1.5/5))

@@

mut_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```
